---
title: "Multiple Solutions GIMME"
author: "Cara Arizmendi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This vignette will cover basic functionality of ms-gimme including how to run ms-gimme, how to avoid potential problems when using ms-gimme, and how to use the solution.tree function to explore the branching of multiple solutions.

Ms-gimme is a modification of gimme (group iterative multiple model estimation), allowing for the recovery of multiple uSEM (unified SEM) solutions with a data driven approach (Beltz and Molenaar, 2016).

## Selecting multiple solutions gimme

To begin, we load a fitted gimme object, with which multiple solutions were found.  This object is available as part of the gimme package and can be accessed as shown below.

```{r}
library(gimme)
load("ms.gimme.obj")
```

Here we show the syntax used to arrive at the fitted object.  To return multiple solutions, ms_allow should be TRUE.  We only show the most relevant arguments to ms-gimme.  We recommend that the default ar=FALSE be kept, as multiple solutions are unlikely to be found when ar=TRUE.  Additionally, the default of subgroup=FALSE should be kept, as ms-gimme does not have the functionality for subgrouping.  The remaining arguments are up to user discretion.  For general gimme usage, we recommend standardize=TRUE; however, this is not the default since standardize=TRUE modifies the data.  We do not recommend that ms_tol be greater than the default, but smaller tolerances can be set.     

```{r, eval=FALSE}
fit <- gimme(data        = beltz2016,
             ar          = FALSE,
             subgroup    = FALSE,
             standardize = TRUE,
             ms_allow    = TRUE,
             ms_tol      = 1e-5)
```

## Exploring output object

### Tree table

The function solution.tree allows for exploration of when paths diverge.  

```{r}
solution.tree(fit, level = "group", cols = c("stage", "pruned"))
```

### Tree plot

A plot of solution divergences is produced when plot.tree=TRUE.

```{r}
solution.tree(fit, level = "group", plot.tree = TRUE)
```


## Exploring output folder

Description of what to expect in output folder.  Explain how individual models are chosen and how users can make other decisions based on the information.

## Deciding between models

We refer users to Beltz and Molenaar (2016) for guidance on model selection.  The AIC is the most widely used fit statistic for model selection and can be accessed in the gimme object as shown below or in the output folder.

```{r}

```


Adriene M. Beltz & Peter C. M. Molenaar (2016) Dealing with Multiple Solutions in Structural Vector Autoregressive Models, Multivariate Behavioral Research, 51:2-3, 357-373, DOI: 10.1080/00273171.2016.1151333
