---
title: "Multiple Solutions GIMME"
author: "Cara Arizmendi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Brief description of ms-gimme with reference to article.  Why someone might want to use ms-gimme.  What the vignette will cover.

## Selecting multiple solutions gimme

We will need some data that either runs in a couple minutes or we don't run this part and have a gimme object already in the package.

```{r}
library(gimme)
load("~/gimme/gimme/data/fit.RData")
```


```{r, eval=FALSE}
fit <- gimme(data        = simData,
             ar          = FALSE,
             plot        = TRUE,
             subgroup    = FALSE,
             paths       = NULL,
             mean_center_mult = FALSE,
             standardize = FALSE,
             groupcutoff = .75,
             subcutoff   = .5,
             diagnos     = FALSE, 
             ms_allow         = TRUE,
             ms_tol           = 1e-5)
```


Discuss important defaults (especially standardize=FALSE).  Describe why someone might want to adjust a default.

## Exploring output object


## Plotting tree of solutions

Reference tree package.

### Tree table

Go into detail about what indicates multiple solutions.  How do we interpret this table?

```{r}
solution.tree(fit, level = "group", cols = c("stage", "pruned"))
```

### Tree plot

Again, how do we interpret this plot?

```{r}
solution.tree(fit, level = "group", plot.tree = TRUE)
```


## Exploring output folder

Description of what to expect in output folder.  Explain how individual models are chosen and how users can make other decisions based on the information.

## Deciding between models

Provide guidance for interpretation and decision making process based on fit indices.


